---
title: "classes vs environment - R Notebook"
output: html_notebook
---

For now envrionmental data was obtained by outputting the sample info from 4a_megaTwinDeepSummariesVclasses with the P class info attached, and using arcGIS to sample the rasters used in the seaPen study. This can be redone with the best available data after Margaret has collected it and the sampling can be performed in R.

#Libraries
```{r}
library(ggplot2)
library(raster)
library(rgdal)
library(tibble)
library(dplyr)
library(ggbiplot)
```
#assemble env var and classes/samples


```{r}
samp.clss<-my.subs %>%
  dplyr::select(SampID, class, level1, level2, level3, level4, level5, level6, level7, level8, level9, level10, level11, level12, level13, level14, level15)


```

# Load Data Env Data (if not already in environment: samp.EnvVar)
```{r}
##If not already loaded you can load the environmental data per sample file that was output in 2b_data_prep_envVarPerSamp.Rmd
# samp.envVar <- read.csv(file.path(outPath, ("samp_envVar.csv")))
```

#Merge class list w samp.EnvVar and add some new SedClasses
```{r}

soft.sed<-c("mud", "clay", "sandy_mud")
coarse.sed<-c("sand", "shellsand","sand_with_shells","sand_with_biomaterial", "clay_spicules")
hard.sed<-c("pebble","coral_rubble", "lophelia_rubble")
hard.solid<-c("bedrock", "boulder", "cobble", "karbonatskorpe","compacted_sediment", "consolidated_clay")


class.env<-samp.clss%>%
  left_join(samp.envVar)%>%
  rowwise()%>%
  dplyr::mutate(softSed=sum(c_across(any_of(soft.sed)), na.rm = T),
         coarseSed=sum(c_across(any_of(coarse.sed)), na.rm = T),
         hardSed=sum(c_across(any_of(hard.sed)), na.rm = T),
         hardSolid=sum(c_across(any_of(hard.solid)), na.rm = T))%>%
  ungroup()
```






#reorder levels of class to be in order

```{r}
#class.ls<-c("P1","P2","P3","P4","P5","P6","P7","P8","P9","P10",
#            "P11","P12","P13","P14","P15","P16","P17","P18","P19","P26",
#            "P20","P21","P22","P23","P24","P25")

class.ls<-c("P1","P2","P3","P4","P5","P6","P7","P8","P9","P10",
            "P11","P12","P13","P14","P15","P16","P17","P18","P19","P20",
            "P21","P22","P23","P24","P25","P26","P27","P28","P29","P30",
            "P31","P32","P33","P34","P35","P36","P37","P38","P39","P40",
            "P41","P42","P43","P44","P45","P46","P47","P48","P49","P50",
            "P51","P52","P53","P54","P55","P56","P57","P58","P59","P60",
            "P61","P62","P63","P64")

class.env$class <- factor(class.env$class, levels = class.ls)

class.env$level1<-as.factor(class.env$level1)
class.env$level2<-as.factor(class.env$level2)
class.env$level3<-as.factor(class.env$level3)
class.env$level4<-as.factor(class.env$level4)
class.env$level5<-as.factor(class.env$level5)
class.env$level6<-as.factor(class.env$level6)
class.env$level7<-as.factor(class.env$level7)
class.env$level8<-as.factor(class.env$level8)
class.env$level9<-as.factor(class.env$level9)
class.env$level10<-as.factor(class.env$level10)
class.env$level11<-as.factor(class.env$level11)
class.env$level12<-as.factor(class.env$level12)
class.env$level13<-as.factor(class.env$level13)
class.env$level14<-as.factor(class.env$level14)
class.env$level15<-as.factor(class.env$level15)

class.env$mean_depth<-class.env$mean_depth*-1

class.env[class.env==-9999]<-NA

summary(class.env)

summary(class.env$class)

colnames(class.env)

```




#Plots

Depth (Bathy)
```{r}
png("U:/Mareano/VIDEOLAB/VIDEO DATA/200m_scale_species_by_sample/Data_Delivery_2021/outputs/Paal class subgroup tables/64subgps_videoEnv/bathy.png",width = 1200, height = 500, units = "px")
ggplot(class.env, aes(x=class, y=mean_depth))+
  geom_boxplot()
dev.off()
```
Temp Av (T_Av)
```{r}
png("U:/Mareano/VIDEOLAB/VIDEO DATA/200m_scale_species_by_sample/Data_Delivery_2021/outputs/Paal class subgroup tables/64subgps_videoEnv/T_Av.png",width = 1200, height = 500, units = "px")
ggplot(class.env, aes(x=class, y=temp_mean..tif))+
  geom_boxplot()
dev.off()
```
Temp Min (T_Min)
```{r}
png("U:/Mareano/VIDEOLAB/VIDEO DATA/200m_scale_species_by_sample/Data_Delivery_2021/outputs/Paal class subgroup tables/64subgps_videoEnv/T_Min.png",width = 1200, height = 500, units = "px")
ggplot(class.env, aes(x=class, y=temp_min..tif))+
  geom_boxplot()
dev.off()
```
Temp SD
```{r}
png("U:/Mareano/VIDEOLAB/VIDEO DATA/200m_scale_species_by_sample/Data_Delivery_2021/outputs/Paal class subgroup tables/64subgps_videoEnv/T_SD.png",width = 1200, height = 500, units = "px")
ggplot(class.env, aes(x=class, y=temp_std..tif))+
  geom_boxplot()
dev.off()
```


Salt Av(Salt_Av)
```{r}
png("U:/Mareano/VIDEOLAB/VIDEO DATA/200m_scale_species_by_sample/Data_Delivery_2021/outputs/Paal class subgroup tables/64subgps_videoEnv/Salt_Av.png",width = 1200, height = 500, units = "px")
ggplot(class.env, aes(x=class, y=salt_mean..tif))+
  geom_boxplot()
dev.off()
```

Salt Min (Salt_Min)
```{r}
png("U:/Mareano/VIDEOLAB/VIDEO DATA/200m_scale_species_by_sample/Data_Delivery_2021/outputs/Paal class subgroup tables/64subgps_videoEnv/Salt_Min.png",width = 1200, height = 500, units = "px")
ggplot(class.env, aes(x=class, y=salt_min..tif))+
  geom_boxplot()
dev.off()
```
Salt SD 
```{r}
png("U:/Mareano/VIDEOLAB/VIDEO DATA/200m_scale_species_by_sample/Data_Delivery_2021/outputs/Paal class subgroup tables/64subgps_videoEnv/Salt_SD.png",width = 1200, height = 500, units = "px")
ggplot(class.env, aes(x=class, y=salt_std..tif))+
  geom_boxplot()
dev.off()
```

Cspd_Av
```{r}
png("U:/Mareano/VIDEOLAB/VIDEO DATA/200m_scale_species_by_sample/Data_Delivery_2021/outputs/Paal class subgroup tables/64subgps_videoEnv/Cspd_Av.png",width = 1200, height = 500, units = "px")
ggplot(class.env, aes(x=class, y=spd_mean..tif))+
  geom_boxplot()
dev.off()
```

Cspd_Mx
```{r}
png("U:/Mareano/VIDEOLAB/VIDEO DATA/200m_scale_species_by_sample/Data_Delivery_2021/outputs/Paal class subgroup tables/64subgps_videoEnv/Cspd_Mx.png",width = 1200, height = 500, units = "px")
ggplot(class.env, aes(x=class, y=spd_max..tif))+
  geom_boxplot()
dev.off()
```
Cspd_Mx
```{r}
png("U:/Mareano/VIDEOLAB/VIDEO DATA/200m_scale_species_by_sample/Data_Delivery_2021/outputs/Paal class subgroup tables/64subgps_videoEnv/Cspd_SD.png",width = 1200, height = 500, units = "px")
ggplot(class.env, aes(x=class, y=spd_std..tif))+
  geom_boxplot()
dev.off()
```

Xcoord
```{r}
png("U:/Mareano/VIDEOLAB/VIDEO DATA/200m_scale_species_by_sample/Data_Delivery_2021/outputs/Paal class subgroup tables/64subgps_videoEnv/Xcoord.png",width = 1200, height = 500, units = "px")
ggplot(class.env, aes(x=class, y=x_coord))+
  geom_boxplot()
dev.off()
```

Ycoord
```{r}
png("U:/Mareano/VIDEOLAB/VIDEO DATA/200m_scale_species_by_sample/Data_Delivery_2021/outputs/Paal class subgroup tables/64subgps_videoEnv/Ycoord.png",width = 1200, height = 500, units = "px")
ggplot(class.env, aes(x=class, y=y_coord))+
  geom_boxplot()
dev.off()
```

# SEDIMENTS
 "boulder"               "cobble"               
[37] "pebble"                "sandy_mud"             "coral_rubble"         
[40] "dead_lophelia"         "live_lophelia"         "sand"                 
[43] "lophelia_rubble"       "bedrock"               "compacted_sediment"   
[46] "mud"                   "clay_(outcropping)"    "shellsand"            
[49] "karbonatskorpe"        "biobush"               "consolidated_clay"    
[52] "clay_spicules"         "gelatinous_snow"       "clay"                 
[55] "sand_with_shells"      "sand_with_biomaterial"

Boulders
```{r}
png("U:/Mareano/VIDEOLAB/VIDEO DATA/200m_scale_species_by_sample/Data_Delivery_2021/outputs/Paal class subgroup tables/64subgps_videoEnv/Sed_Bould.png",width = 1200, height = 500, units = "px")
ggplot(class.env, aes(x=class, y=boulder))+
  geom_boxplot()
dev.off()
```
Cobbles
```{r}
png("U:/Mareano/VIDEOLAB/VIDEO DATA/200m_scale_species_by_sample/Data_Delivery_2021/outputs/Paal class subgroup tables/64subgps_videoEnv/Sed_Cobb.png",width = 1200, height = 500, units = "px")
ggplot(class.env, aes(x=class, y=cobble))+
  geom_boxplot()
dev.off()
``` 
Pebbles
```{r}
png("U:/Mareano/VIDEOLAB/VIDEO DATA/200m_scale_species_by_sample/Data_Delivery_2021/outputs/Paal class subgroup tables/64subgps_videoEnv/Sed_Pebb.png",width = 1200, height = 500, units = "px")
ggplot(class.env, aes(x=class, y=pebble))+
  geom_boxplot()
dev.off()
``` 
Sandy_mud
```{r}
png("U:/Mareano/VIDEOLAB/VIDEO DATA/200m_scale_species_by_sample/Data_Delivery_2021/outputs/Paal class subgroup tables/64subgps_videoEnv/Sed_SandyMud.png",width = 1200, height = 500, units = "px")
ggplot(class.env, aes(x=class, y=sandy_mud))+
  geom_boxplot()
dev.off()
``` 

SedSand (Sed_Sand)
```{r}
png("U:/Mareano/VIDEOLAB/VIDEO DATA/200m_scale_species_by_sample/Data_Delivery_2021/outputs/Paal class subgroup tables/64subgps_videoEnv/Sed_Sand.png",width = 1200, height = 500, units = "px")
ggplot(class.env, aes(x=class, y=sand))+
  geom_boxplot()
dev.off()
```

SedMud
```{r}
png("U:/Mareano/VIDEOLAB/VIDEO DATA/200m_scale_species_by_sample/Data_Delivery_2021/outputs/Paal class subgroup tables/64subgps_videoEnv/Sed_Mud.png",width = 1200, height = 500, units = "px")
ggplot(class.env, aes(x=class, y=mud))+
  geom_boxplot()
dev.off()
```
Bedrock
```{r}
png("U:/Mareano/VIDEOLAB/VIDEO DATA/200m_scale_species_by_sample/Data_Delivery_2021/outputs/Paal class subgroup tables/64subgps_videoEnv/Sed_Bedrock.png",width = 1200, height = 500, units = "px")
ggplot(class.env, aes(x=class, y=bedrock))+
  geom_boxplot()
dev.off()
```
Compacted Sediment
```{r}
png("U:/Mareano/VIDEOLAB/VIDEO DATA/200m_scale_species_by_sample/Data_Delivery_2021/outputs/Paal class subgroup tables/64subgps_videoEnv/Sed_CompSed.png",width = 1200, height = 500, units = "px")
ggplot(class.env, aes(x=class, y=compacted_sediment))+
  geom_boxplot()
dev.off()
```
shellsand
```{r}
png("U:/Mareano/VIDEOLAB/VIDEO DATA/200m_scale_species_by_sample/Data_Delivery_2021/outputs/Paal class subgroup tables/64subgps_videoEnv/Sed_Shellsand.png",width = 1200, height = 500, units = "px")
ggplot(class.env, aes(x=class, y=shellsand))+
  geom_boxplot()
dev.off()
```
clay
```{r}
png("U:/Mareano/VIDEOLAB/VIDEO DATA/200m_scale_species_by_sample/Data_Delivery_2021/outputs/Paal class subgroup tables/64subgps_videoEnv/Sed_Clay.png",width = 1200, height = 500, units = "px")
ggplot(class.env, aes(x=class, y=clay))+
  geom_boxplot()
dev.off()
```
consolidated_clay
```{r}
png("U:/Mareano/VIDEOLAB/VIDEO DATA/200m_scale_species_by_sample/Data_Delivery_2021/outputs/Paal class subgroup tables/64subgps_videoEnv/Sed_ConsolClay.png",width = 1200, height = 500, units = "px")
ggplot(class.env, aes(x=class, y=consolidated_clay))+
  geom_boxplot()
dev.off()
```
clay outcropping
```{r}
png("U:/Mareano/VIDEOLAB/VIDEO DATA/200m_scale_species_by_sample/Data_Delivery_2021/outputs/Paal class subgroup tables/64subgps_videoEnv/Sed_ClayOutcrop.png",width = 1200, height = 500, units = "px")
ggplot(class.env, aes(x=class, y="clay_(outcropping)"))+
  geom_boxplot()
dev.off()
```

karbonatskorpe
```{r}
png("U:/Mareano/VIDEOLAB/VIDEO DATA/200m_scale_species_by_sample/Data_Delivery_2021/outputs/Paal class subgroup tables/64subgps_videoEnv/Sed_KarbonatSk.png",width = 1200, height = 500, units = "px")
ggplot(class.env, aes(x=class, y=karbonatskorpe))+
  geom_boxplot()
dev.off()
```
biobush
```{r}
png("U:/Mareano/VIDEOLAB/VIDEO DATA/200m_scale_species_by_sample/Data_Delivery_2021/outputs/Paal class subgroup tables/64subgps_videoEnv/Sed_Biobush.png",width = 1200, height = 500, units = "px")
ggplot(class.env, aes(x=class, y=biobush))+
  geom_boxplot()
dev.off()
```
clay_spicules
```{r}
png("U:/Mareano/VIDEOLAB/VIDEO DATA/200m_scale_species_by_sample/Data_Delivery_2021/outputs/Paal class subgroup tables/64subgps_videoEnv/Sed_ClaySpicules.png",width = 1200, height = 500, units = "px")
ggplot(class.env, aes(x=class, y=clay_spicules))+
  geom_boxplot()
dev.off()
```
gelatinous_snow
```{r}
png("U:/Mareano/VIDEOLAB/VIDEO DATA/200m_scale_species_by_sample/Data_Delivery_2021/outputs/Paal class subgroup tables/64subgps_videoEnv/Sed_GelSnow.png",width = 1200, height = 500, units = "px")
ggplot(class.env, aes(x=class, y=gelatinous_snow))+
  geom_boxplot()
dev.off()
```
sand with shells
```{r}
png("U:/Mareano/VIDEOLAB/VIDEO DATA/200m_scale_species_by_sample/Data_Delivery_2021/outputs/Paal class subgroup tables/64subgps_videoEnv/Sed_SandWShells.png",width = 1200, height = 500, units = "px")
ggplot(class.env, aes(x=class, y=sand_with_shells))+
  geom_boxplot()
dev.off()
```
sand with biomaterial
```{r}
png("U:/Mareano/VIDEOLAB/VIDEO DATA/200m_scale_species_by_sample/Data_Delivery_2021/outputs/Paal class subgroup tables/64subgps_videoEnv/Sed_SandWBiomateral.png",width = 1200, height = 500, units = "px")
ggplot(class.env, aes(x=class, y=sand_with_biomaterial))+
  geom_boxplot()
dev.off()
```






#Coral/Lophelia

Coral Rubble
```{r}
png("U:/Mareano/VIDEOLAB/VIDEO DATA/200m_scale_species_by_sample/Data_Delivery_2021/outputs/Paal class subgroup tables/64subgps_videoEnv/Sed_CoralRubb.png",width = 1200, height = 500, units = "px")
ggplot(class.env, aes(x=class, y=coral_rubble))+
  geom_boxplot()
dev.off()
```
Lophelia Rubble
```{r}
png("U:/Mareano/VIDEOLAB/VIDEO DATA/200m_scale_species_by_sample/Data_Delivery_2021/outputs/Paal class subgroup tables/64subgps_videoEnv/Sed_LopheliaRubb.png",width = 1200, height = 500, units = "px")
ggplot(class.env, aes(x=class, y=lophelia_rubble))+
  geom_boxplot()
dev.off()
```
Live Lophelia
```{r}
png("U:/Mareano/VIDEOLAB/VIDEO DATA/200m_scale_species_by_sample/Data_Delivery_2021/outputs/Paal class subgroup tables/64subgps_videoEnv/Sed_LiveLophelia.png",width = 1200, height = 500, units = "px")
ggplot(class.env, aes(x=class, y=live_lophelia))+
  geom_boxplot()
dev.off()
```
Live Lophelia
```{r}
png("U:/Mareano/VIDEOLAB/VIDEO DATA/200m_scale_species_by_sample/Data_Delivery_2021/outputs/Paal class subgroup tables/64subgps_videoEnv/Sed_LiveLophelia.png",width = 1200, height = 500, units = "px")
ggplot(class.env, aes(x=class, y=live_lophelia))+
  geom_boxplot()
dev.off()
```

Dead Lophelia
```{r}
png("U:/Mareano/VIDEOLAB/VIDEO DATA/200m_scale_species_by_sample/Data_Delivery_2021/outputs/Paal class subgroup tables/64subgps_videoEnv/Sed_DeadLophelia.png",width = 1200, height = 500, units = "px")
ggplot(class.env, aes(x=class, y=dead_lophelia))+
  geom_boxplot()
dev.off()
```
#Bex new sedClasses
soft sed
```{r}
png("U:/Mareano/VIDEOLAB/VIDEO DATA/200m_scale_species_by_sample/Data_Delivery_2021/outputs/Paal class subgroup tables/64subgps_videoEnv/Sed_SOFT.png",width = 1200, height = 500, units = "px")
ggplot(class.env, aes(x=class, y=softSed))+
  geom_boxplot()
dev.off()
```
coarse sed
```{r}
png("U:/Mareano/VIDEOLAB/VIDEO DATA/200m_scale_species_by_sample/Data_Delivery_2021/outputs/Paal class subgroup tables/64subgps_videoEnv/Sed_COARSE.png",width = 1200, height = 500, units = "px")
ggplot(class.env, aes(x=class, y=coarseSed))+
  geom_boxplot()
dev.off()
```
hard sed
```{r}
png("U:/Mareano/VIDEOLAB/VIDEO DATA/200m_scale_species_by_sample/Data_Delivery_2021/outputs/Paal class subgroup tables/64subgps_videoEnv/Sed_HARD_SED.png",width = 1200, height = 500, units = "px")
ggplot(class.env, aes(x=class, y=hardSed))+
  geom_boxplot()
dev.off()
```
hard solid
```{r}
png("U:/Mareano/VIDEOLAB/VIDEO DATA/200m_scale_species_by_sample/Data_Delivery_2021/outputs/Paal class subgroup tables/64subgps_videoEnv/Sed_HARD_SOLID.png",width = 1200, height = 500, units = "px")
ggplot(class.env, aes(x=class, y=hardSolid))+
  geom_boxplot()
dev.off()
```
