---
title: "4bb - modellingEnvChecks vs Class - R Notebook"
output: html_notebook
---

If you have run all previous environmental checks with different environmental data from what is used in the models, it can be helpful to get an overview of how the classes align with the modelling environmental variables to see which classess might react best/worst in models/ which might be worth merging on the grounds of not enough resolution in modelling variables to tell them apart. This is where you can load those envVar and start exploring how the classes interact with them.

#Libraries
```{r}
library(dplyr)
library(tidyr)
library(raster)
```

#Data loading
Need:

- a stack of envrionmental variable rasters used in the modelling, and 
- the shapefile from 4a that holds the latest sample, class, x/y info, most likely my.shp.s 

Check my.shp.s

```{r}
plot(my.shp.s)
summary(my.shp.s@data)

```


Load any rasters to sample

```{r}
rasPath<-"U:/Mareano/VIDEOLAB/VIDEO DATA/200m_scale_species_by_sample/Data_Delivery_2021/envStack/"
ras.files.names<-list.files(rasPath)
ras.files<-list.files(rasPath,full.names = T)
ras.stk<-stack(ras.files)
names(ras.stk)<-unlist(ras.files.names)
```


#sample rasters and convert to dataframe
```{r}
smpClassEnvMod.shp<-raster::extract(ras.stk, my.shp.s, sp=T)
smpClassEnvMod.o<-as.data.frame(smpClassEnvMod.shp)#%>%
  #dplyr::select(-c("coords.x1","coords.x2"))
write.csv(smpClassEnvMod.o, file = file.path(outPath, ("samp_Class_envVar_Modelling_SubTmpMrg.csv")))
```

