---
title: "betapart - R Notebook"
output: html_notebook
---

Checking what we can do with some beta diversity metrics when assessing potential biotopes. Requires that you have already done a twinspan analysis (making the object tw_1) and have decided on some provisional biotopes/classes. Currently developed to run after you have done some megaTwinDeepSummaries and have object "piv.av" (pseudospecies averaged per class with class columns, species rows, and average pseudospecies values in th matrix)

```{r}
library(dplyr)
library(tibble)
library(tidyr)
library(betapart)

```

## format data
needs to be in the format of a matrix (x) codifying the presence (1) or absence (0) of m species (columns) in n sites (rows).
We want the columns sorted to avoid a disorganised table at the end.

```{r}
## if you need to sort, use this code - listshould be in the order you need
# class.ls<-c("P1","P2","P3","P4","P5","P6","P7","P8","P9","P10",
#                 "P11","P12","P13","P14","P15","P16","P17","P18","P19","P26",
#                 "P20","P21","P22","P23","P24","P25")
# 
# beta.in <-piv.av %>%
#   as.data.frame()%>%
#   select(class.ls)

beta.in<-piv.av

beta.in[beta.in>0]=1
beta.in<-t(beta.in)
```

## run betapart
```{r}
beta.out<-betapart.core(beta.in)
beta.share<-beta.out[["shared"]]
```

## find union of samples
total number of taxa for each pair
```{r}
#rowSumes
rsum.beta.in<-rowSums(beta.in) %>%
  as.data.frame()

#for relabelling lost rownames
name.lookup <- data.frame(class=rownames(beta.in),
                          V =c(1:length(rownames(beta.in))))

#pairwise sums of rowsums
pair.sums.beta.in<-t(combn(nrow(rsum.beta.in), 2))
pair.sums.beta.calc <- apply(pair.sums.beta.in, 1, function(x) sum(rsum.beta.in[x[1],] + rsum.beta.in[x[2],]))
pair.sums.beta.out<-cbind(pair.sums.beta.in, pair.sums.beta.calc)

#re-add class names
pair.sums.beta.out.labelled<-pair.sums.beta.out%>%
  as.data.frame()%>%
  inner_join(name.lookup, by =c('V1'='V'))%>%
  inner_join(name.lookup, by =c('V2'='V'))%>%
  select(class.x, class.y, pair.sums.beta.calc)

unions<-pair.sums.beta.out.labelled %>%
  as.data.frame()%>%
  pivot_wider(names_from = class.y,
              values_from= pair.sums.beta.calc)%>%
  column_to_rownames(var="class.x")

#make match shape of beta.share (add P1 column and P25 row both filled with NA)
unions<-unions %>%
  mutate(SubTmpMrgA=NA)%>%              ##### MAKE SURE HAS CORRECT NAME
  select(SubTmpMrgA, everything())%>%   ##### MAKE SURE HAS CORRECT NAME
  add_row()%>%
#  `rownames<-`(class.ls)
  `rownames<-`(classMerge.ls)           ##### MAKE SURE HAS CORRECT LIST



```



## find proportion shared
```{r}

beta.share.out<-beta.share/unions#/(beta.out[["shared"]]+beta.out[["not.shared"]])

round_df <- function(x, digits) {
    # round all numeric variables
    # x: data frame 
    # digits: number of digits to round
    numeric_columns <- sapply(x, mode) == 'numeric'
    x[numeric_columns] <-  round(x[numeric_columns], digits)
    x
}

beta.share.out<-round_df(beta.share.out, 2)
```

## write to csv
```{r}
write.csv(beta.share.out, 
          file = file.path(outPath, ("SubTmpMrg_beta_sharePropOfUnion_TOPCOPY.csv"))) #change label
```

