---
title: "Bex explorations - R Notebook"
output: html_notebook
---
Will be copied into data explore later if any of this is useful

Assumes you have run data_analysis_twinspan beforehand and have tw_1 loaded in the environment.

# input table
```{r}
in.data<-trimmedwide %>%
  filter(new_filter) %>%
  filter(sample_info$VL[match(SampID, sample_info$SampID)]%in%good) %>%
  select(-SampID) %>%
  select(which(!is.na(colSums(.))))
```




#twintable
```{r}
tw_1.xl<-twintable(tw_1)

capture.output(twintable(tw_1), file = file.path(outPath,"twintable_tw1.csv"))

lapply(tw_1.xl, function(x) write.table( data.frame(x), file.path(outPath,"twintable_tw1.csv")  , append= T, sep=',' ))

write.csv(tw_1.xl[["table"]],file.path(outPath,"twintable_tw1.csv"))
```


#dendrogram
samples
```{r}
plot(tw_1, "quadrat")

pdf(file.path(outPath,"miniDendro_tw1.pdf"), width=20, height=10)

# Do some plotting
plot(tw_1,"quadrat")

# Close the PDF file's associated graphics device (necessary to finalize the output)
dev.off()
```

species
```{r}
plot(tw_1,"species")

pdf(file.path(outPath,"miniDendro_tw1_spp.pdf"), width=20, height=10)

# Do some plotting
plot(tw_1,"species")

# Close the PDF file's associated graphics device (necessary to finalize the output)
dev.off()
```

detailed pdfs (just massive so can zoom in)
```{r}
pdf(file.path(outPath,"giantDendro_tw1_triangle.pdf"), width=100, height=10)
par(cex=0.1)
# Do some plotting
plot(as.dendrogram(tw_1,"quadrat"),type="triangle")

# Close the PDF file's associated graphics device (necessary to finalize the output)
dev.off()


pdf(file.path(outPath,"giantDendro_tw1_triangle_spp.pdf"), width=100, height=15)
par(cex=1)
# Do some plotting
plot(as.dendrogram(tw_1,"species"),type="triangle")

# Close the PDF file's associated graphics device (necessary to finalize the output)
dev.off()


pdf(file.path(outPath,"giantDendro_tw1.pdf"), width=100, height=10)
par(cex=0.1)
# Do some plotting
plot(as.dendrogram(tw_1,"quadrat"))

# Close the PDF file's associated graphics device (necessary to finalize the output)
dev.off()


pdf(file.path(outPath,"giantDendro_tw1_spp.pdf"), width=100, height=10)
par(cex=1)
# Do some plotting
plot(as.dendrogram(tw_1,"species"))

# Close the PDF file's associated graphics device (necessary to finalize the output)
dev.off()
```
#Understanding indicators
subset to eucratea samples to cross check with rules
```{r}
eu.samps<-subset(trimmedwide, eucratealoricata>0)
```



-----------



#deeper twinspan
Currently stops at 6 levels, maximu is 15, so could try 15, but make sure there are at least 5 samples in each terminus? Then can collapse back up

```{r}
cutlevelsPaal = c(0,1,2,5,10,20)

tw_1_15lev <- trimmedwide %>%
  filter(new_filter) %>%
  filter(sample_info$VL[match(SampID, sample_info$SampID)]%in%good) %>%
  select(-SampID) %>%
  select(which(!is.na(colSums(.)))) %>%
  #mutate_all(sqrt)%>%
  twinspan(cutlevels = cutlevelsPaal, levmax=15, groupmin=5)
```
#dendrogram
samples
```{r}
plot(tw_1_15lev, "quadrat")

pdf(file.path(outPath,"miniDendro_tw1_15lev.pdf"), width=80, height=10)
par(cex=0.5)
# Do some plotting
plot(tw_1_15lev,"quadrat")

# Close the PDF file's associated graphics device (necessary to finalize the output)
dev.off()
```

species
```{r}
plot(tw_1_15lev,"species")

pdf(file.path(outPath,"miniDendro_tw1_15lev_spp.pdf"), width=20, height=10)

# Do some plotting
plot(tw_1_15lev,"species")

# Close the PDF file's associated graphics device (necessary to finalize the output)
dev.off()
```

detailed pdfs (just massive so can zoom in)
```{r}
pdf(file.path(outPath,"giantDendro_tw1_15lev_triangle.pdf"), width=100, height=10)
par(cex=0.1)
# Do some plotting
plot(as.dendrogram(tw_1_15lev,"quadrat"),type="triangle")

# Close the PDF file's associated graphics device (necessary to finalize the output)
dev.off()


pdf(file.path(outPath,"giantDendro_tw1_15lev_triangle_spp.pdf"), width=100, height=15)
par(cex=1)
# Do some plotting
plot(as.dendrogram(tw_1_15lev,"species"),type="triangle")

# Close the PDF file's associated graphics device (necessary to finalize the output)
dev.off()


pdf(file.path(outPath,"giantDendro_tw1_15lev.pdf"), width=100, height=10)
par(cex=0.1)
# Do some plotting
plot(as.dendrogram(tw_1_15lev,"quadrat"))

# Close the PDF file's associated graphics device (necessary to finalize the output)
dev.off()


pdf(file.path(outPath,"giantDendro_tw1_15lev_spp.pdf"), width=100, height=10)
par(cex=1)
# Do some plotting
plot(as.dendrogram(tw_1_15lev,"species"))

# Close the PDF file's associated graphics device (necessary to finalize the output)
dev.off()
```
