---
title: "Bex's explorations - R Notebook"
output: html_notebook
---
Will be copied into data explore later if any of this is useful

Assumes you have run data_analysis_twinspan beforehand and have tw_1 loaded in the environment.

#additional libraries
```{r}
library(RColorBrewer)
library(heatmaply)
# library(viridis)
# library(wesanderson)
```




# input table
Just in case you want to check something in it
```{r}
in.data<-coredata %>%
  select(which(!is.na(colSums(.))))
```



## store sample labels
use to switch with q labels

- GENO HAS AMENDED IN STEP PRIOR TO RUNNING TWINSPAN
```{r}
# samples <- trimmedwide %>%
#   data.frame() %>%
#   filter(new_filter) %>%
#   filter(sample_info$VL[match(SampID, sample_info$SampID)]%in%good) %>%
#   select(SampID)
# 
# tw_1samp<-tw_1
# 
# tw_1samp[["quadrat"]][["labels"]]<-as.vector(samples[["SampID"]])
  
```






#twintable
Takes ages to run (e.g. 30mins) and the output is near unusable, suggest doing it with smaller chunks only else we need to figure out how to digest this output better/ recreate the output
```{r}
#tw_1.xl<-twintable(tw_1)
# head(twintable(tw_1))
# capture.output(twintable(tw_1), file = file.path(outPath,"twintable_tw1.csv"))
# 
# lapply(tw_1.xl, function(x) write.table( data.frame(x), file.path(outPath,"twintable_tw1.csv")  , append= T, sep=',' ))
# 
# write.csv(tw_1.xl[["table"]],file.path(outPath,"twintable_tw1.csv"))

```

## example smaller chunk twintable

```{r}
twintable(tw_1samp, subset = cut(tw_1samp, 1) == 3, goodspecies = "both")
```



#dendrograms

## A. 

TWINSPAN preferred output with nodes labelled as splits, and groups at the terminus with number of samples. Best output as pdf where you can adjust the size of the paper and then zoom in. May need to adjust widths when rerun.

samples
```{r}
plot(tw_1samp, "quadrat")

pdf(file.path(outPath,"miniDendro_tw1.pdf"), width=200, height=10) #adjust width as necessary
par(cex=0.7)
# Do some plotting
plot(tw_1samp,"quadrat")

# Close the PDF file's associated graphics device (necessary to finalize the output)
dev.off()
```

species
```{r}
plot(tw_1samp,"species")

pdf(file.path(outPath,"miniDendro_tw1_spp.pdf"), width=50, height=10)

# Do some plotting
plot(tw_1samp,"species")

# Close the PDF file's associated graphics device (necessary to finalize the output)
dev.off()
```

## B.

The as.dendrogram plots do not label the nodes with splits, but they do have the quadrats (internal sample numbers) at the terminus (or the species in the species dendrogram). Again best output with large pdf pages (1 m wide, lol) so that you can zoom in or print out.


detailed pdfs (just massive so can zoom in)
```{r}
pdf(file.path(outPath,"giantDendro_tw1_triangle_sampleID.pdf"), width=100, height=10)
par(cex=0.1, mar=c(10,0,0,0))
# Do some plotting
plot(as.dendrogram(tw_1samp,"quadrat"),type="triangle")

# Close the PDF file's associated graphics device (necessary to finalize the output)
dev.off()


pdf(file.path(outPath,"giantDendro_tw1_triangle_spp.pdf"), width=100, height=15)
par(cex=1,mar=c(10,0,0,0))
# Do some plotting
plot(as.dendrogram(tw_1,"species"),type="triangle")

# Close the PDF file's associated graphics device (necessary to finalize the output)
dev.off()
plot(as.dendrogram(tw_1,"species"),type="triangle")

# pdf(file.path(outPath,"giantDendro_tw1.pdf"), width=100, height=10)
# par(cex=0.1)
# # Do some plotting
# plot(as.dendrogram(tw_1,"quadrat"))
# 
# # Close the PDF file's associated graphics device (necessary to finalize the output)
# dev.off()
# 
# 
# pdf(file.path(outPath,"giantDendro_tw1_spp.pdf"), width=100, height=10)
# par(cex=1)
# # Do some plotting
# plot(as.dendrogram(tw_1,"species"))
# 
# # Close the PDF file's associated graphics device (necessary to finalize the output)
# dev.off()
```
# Developing better plots
Going to start by subsetting to a cut to see what we can do with smaller subsets

## Image

```{r}
colrmp<-rev(brewer.pal(6,"OrRd"))
image(tw_1, reorder=T,col=colrmp)

# better output to a giant pdf where you can see all quadrats listed
#par(cexCol=0.2, cexRow=10)
#par(cex=0.2)
pdf(file.path(outPath,"giant_heatmap_tw1.pdf"), width=50, height=50)
image(tw_1, reorder=T, col=colrmp, margins=c(10,10))
dev.off()
```
## subset heatmap
UNDER DEVELOPMENT
```{r}
cucu<- cut(tw_1, 1)

colrmp<-rev(brewer.pal(6,"OrRd"))
image(cucu, reorder=T, col=colrmp)
```




## exploring making an interactive heatmap
UNDER DEVELOPMENT
looking into heatmaply
```{r}
mat<-twin2mat(tw_1)
```













