---
title: "install_twinspan"
author: "GGM & RR"
date: "8 6 2021"
output: html_document
---

```{r global-options, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, fig.path='Figs/', warning=FALSE, message=FALSE)
```

Geno put this together originally, I (RR) have just updated and further annotated to ensure I can install both versions of TWINSPAN on my machine too and to remember more about how things work. The test for each version of R based TWINSPAN is also included. 

Jarikosa version   
- doesn't need executable, can do various explorations including dendrograms with outputs 

Zdealveindy version   
- needs executable (so R/Rstudio must be run as admin on HI machines), can give output in original format, runs a modified version of TWINSPAN ie. While in the original TWINSPAN, at each level of the division each cluster is divided into two clusters (unless the cluster contains too few samples), in the modified TWINSPAN only the most compositionally heterogeneous cluster is divided into two clusters  


I think, for MAREANO, we should focus on the jarioksa version as there are more options for exploring the output and I dont think we are particularly in need of the 'modified TWINSPAN' that the zdalveindy version was developed to support. Consequently the jarioksa code comes first in this notebook with the zdalveindy version details at the end in case they are needed.


##JARIOKSA TWINSPAN
 
(Can plot dendrograms and use across platforms but can't use modified algorithm)  

Two-Way Indicator Analysis (TWINSPAN) is a divisive classification method that works by splitting first Correspondence Analysis into two classes, and then recursively working with each split subset. The current function is based on and uses much of the original FORTRAN code of the original TWINSPAN (Hill 1979). twinspan is the main function of this package, but it works silently and prints very little information: you must use separate support functions to extract various aspects of the result.

#install
```{r install-jari}
#install.packages ('devtools') # if you don't have it installed yet
library(devtools)
devtools::install_github("jarioksa/twinspan") 
```


#libraries
```{r libs-jari}
library(twinspan)
library(vegan) # for data
```

#test run
https://github.com/jarioksa/twinspan

arguments: 

`twinspan(x, cutlevels = c(0, 2, 5, 10, 20), indmax = 7, groupmin = 5,
  levmax = 6, lind, lwgt, noind)`
  
cutlevels	 
Cut levels used to split quantitative data into binary pseudospecies. Max of 9 cutlevels can be used.

indmax	 
Maximum number of indicators for division (15 or less).

groupmin	 
Minimum group size for division (2 or larger).

levmax	 
Maximum depth of levels of divisions (15 or less).

lind	 
Weights for levels of pseudospecies. For example indicator potentials c(1, 0, 0,1, 0) signify that pseudospecies at levels 1 and 4 can be used as indicators, but that those at other levels cannot. In the default case, all species are available.

lwgt	 
Weights for the levels of pseudospecies. For example weights c(1, 2, 2, 2) signify that pseudospecies corresponding to 3 higher cut levels are to be given twice the weight of pseudospecies at the lowest level.

noind	 
Numbers (indices) of species that you wish to omit from list of potential indicators. Species omitted from this list are used in the calculation, but cannot appear as indicators.

```{r runTwinspJ}
#run with default options
data(varespec)
head(varespec)
twJ <- twinspan(varespec)

```

# explore

To see data transfomation (pseudospecies):
```{r transforms}
twindat <- twinsform(varespec)
colnames(twindat)
```

see classification
```{r classTwinspJ}
summary(twJ)
```

to extract the classification of each quadrat with cut:
```{r perQuarat}
 cut(twJ)
  cut(twJ, level=2) # use classification at second level
```
to predict the membership of quadrats based on pseudospecies and threshold scores
```{r predMemb}
predict(twJ, level=2)
```
classification of species (not pseudospecies - TWINSPAN = two way - samples and species)
```{r classSppTwinspJ}
summary(twJ, "species")
#cut(tw, "species") #for cutting speces classification
```
tabulate: The strings of 0 and 1 in front of the species name and above quadrat name (or number) give the steps of division. The numeric values in the table are the pseudospecies values of the analysis.
```{r tabTwinspJ}
twintable(twJ)

```
dendrogram - quadrats i.e. how quadrats group according to species - primary biotope method
```{r dendroQuadTwinspJ}
plot(as.dendrogram(twJ, "quadrat"), type = "triangle")
```

dendrogram - species i.e. how species group according to quadrats
```{r dendroSpTwinspJ}
plot(as.dendrogram(twJ, "species"), type = "triangle")
```
***
## ZDEALVEINDY TWINSPAN
  
NB RUN AS ADMIN (on HI machines the zdealveindy twinspan run won't work if run as user as there is an executable in the library)  

(can run modified TWINSPAN but only works in windows and can't plot dendrograms) 

# install
```{r installTwinspZ}
#install.packages ('devtools') # if you don't have it installed yet
library(devtools)
devtools::install_github("nsj3/riojaExtra")
devtools::install_github("zdealveindy/twinspanR")
```


# libraries
```{r libsTwinspZ}
detach(package:twinspan,unload=TRUE)
library (twinspanR)
library (vegan)
```


# test run
https://www.davidzeleny.net/anadat-r/doku.php/en:hier-divisive_r  

NB If you have this error when you try and run twinspan then it is probably because you need to run as administrator to get past it...  
`Error in riojaExtra::write.CEP(com, fName = "tw.cc!") : Cannot open file tw.cc!`  

Modified TWINSPAN on traditional Ellenberg's Danube meadow dataset, projected on DCA and compared with original classification into three vegetation types made by tabular sorting:

nb seems to report a stack imbalance sometimes which may need to be reported to david zeleny if it persists and is reproducible.
```{r runTwinspZ}
data (danube)
res <- twinspan (danube$spe, modif = TRUE, clusters = 4)
```
project modified twinspan on DCA ordination diagram and compare to original classification
```{r}
k <- cut (res)
dca <- decorana (danube$spe)
par (mfrow = c(1,2))
ordiplot (dca, type = 'n', display = 'si', main = 'Modified TWINSPAN')
points (dca, col = k)
for (i in c(1,2,4)) ordihull (dca, groups = k, show.group = i, col = i,
                              draw = 'polygon', label = TRUE)
ordiplot (dca, type = 'n', display = 'si', main = 'Original assignment\n (Ellenberg 1954)')
points (dca, col = danube$env$veg.type)
for (i in c(1:3)) ordihull (dca, groups = danube$env$veg.type, 
                            show.group = unique (danube$env$veg.type)[i], col = i,
                            draw = 'polygon', label = TRUE)


```

capture console output from TWINSPAN software
```{r outputTwispZ}
## To capture the console output of twinspan.exe into R object, use the following:
## Not run: 
out <- capture.output (twZ <- twinspan (danube$spe, show.output.on.console = T))
summary (twZ)           # returns summary of twinspan algorithm
cat (out, sep = '\n')  # prints the captured output
write.table (out, file = 'out.txt', quot = F, row.names = F) # writes output to 'out.txt' file

## End(Not run)
```

