---
title: "megaTwin deep summaries - R Notebook"
output: html_notebook
---
As even splitting into smaller twintables can result in some being too big to digest, we want to try and summarise groups at different levels for their average pseudospp., n, max pseudospp? to be better able to compare sub-branches.

#library
```{r}
library(twinspan)
library(dplyr, quietly = TRUE)
library(tibble)
library(Matrix)
library(rgdal)
```

# Load environmental data for later

```{r}
envK <- read.csv(file.path(dataPath,"KF50_samples_env.csv"))
envR <- read.csv(file.path(dataPath,"RF50_samples_env.csv"))

colnames(envK)
colnames(envR)
```

# Create matrix
we need the twintable matrix to work with, which we can get with twin2mat and this needs to be merged with res
```{r}
tw_kr.df<-rownames_to_column(as.data.frame(twin2mat(tw_kr)))


#make res again if need be
resKR<-data.frame(SampID = sites.sel, level1 = NA, level2= NA, level3 = NA, level4 = NA, level5 = NA, level6 = NA, level7 = NA, level8= NA, level9 = NA, level10 = NA, level11 = NA, level12 = NA, level13 = NA, level14= NA, level15 = NA)
 
 for(i in 1:15){
   resKR[,i+1]<-cut(tw_kr, level = i)
 }
 
 f = function(x) length(unique(x))
 lapply(resKR, FUN = f)
 
#join together
res.matKR<-left_join(resKR, tw_kr.df, by = c("sites.sel"="rowname"))


rownames(res.matKR) <- res.matKR[,1]
```
# merge matrix with selected env var
```{r}
#select var
envK.sel<-envK %>%
  select(SampID,
           KF50_bathy_fmn3,
           KF50_slope3_fmn3,
           KF50_Sv800_SALTmin_fmn3,
           KF50_Sv800_TEMPmean_fmn3,
           KF50_Sv800_SPDmax_fmn3,
           sedClass50)
colnames(envK.sel)<-c("SampID",
           "F50_bathy_fmn3",
           "F50_slope3_fmn3",
           "F50_Sv800_SALTmin_fmn3",
           "F50_Sv800_TEMPmean_fmn3",
           "F50_Sv800_SPDmax_fmn3",
           "sedClass50")

envR.sel<-envR %>%
  select(SampID,
           KF50_bathy_fmn3,
           RF50_slope3_fmn3,
           RF50_Sv800_SALTmin_fmn3,
           RF50_Sv800_TEMPmean_fmn3,
           RF50_Sv800_SPDmax_fmn3,
           sedClass50)
colnames(envR.sel)<-c("SampID",
           "F50_bathy_fmn3",
           "F50_slope3_fmn3",
           "F50_Sv800_SALTmin_fmn3",
           "F50_Sv800_TEMPmean_fmn3",
           "F50_Sv800_SPDmax_fmn3",
           "sedClass50")

env<-envK.sel %>%
  bind_rows(envR.sel)

# get positional info
pos<-sample_info %>%
  select(SampID,
         x_coord,
         y_coord)


#merge with dataset

res.matKR<- res.matKR %>%
  left_join(., env, by=c("sites.sel"="SampID"))



rownames(res.matKR) <- res.matKR[,1]
```


# Subset to group at given cut level

```{r}
# Set number of classes needed
n<-7

# Make reclassification table
reclassify <- data.frame(level=c(1,4,4,4,4,3,3),
                         class=c(3,18,19,16,17,11,10),
                         new_class=sprintf("KR%s",seq(1:n)))


res.matKRREF<-res.matKR


# Reclassify
for(i in 1:n){
  res.matKR[getsamplesingrp(res.matKRREF,reclassify$level[i],reclassify$class[i]),"class"]<-as.character(reclassify$new_class[i])
}

table(res.matKR$class)

my.subs<-res.matKR
#my.subs<-subset(res.mat,level8 ==298) #change as needed

```
# output classified shapefile
```{r}
my.subs<-my.subs%>%
  left_join(pos, by=c("sites.sel"="SampID"))


my.subs.shp <- SpatialPointsDataFrame(coords = cbind(my.subs$x_coord,
                                                     my.subs$y_coord),
                                      proj4string = CRS("+init=epsg:32633"),
                                      data =  my.subs)

writeOGR(my.subs.shp, dsn = file.path(outPath, "Spatial"), "KR_classified", driver = "ESRI Shapefile", overwrite_layer = TRUE)

```


# pivot at given cut level

Make table of metadata (parent groups, n sub groups at level 15, and n samples per group)

```{r}
#parent groups
piv.meta.a<-my.subs %>%
  group_by(class)%>% # change to level you want to see groups at
  summarise(across(c(2:15),mean)) %>%
  t()%>%
  `colnames<-`(.[1, ])%>%
  .[-1, ]

# n lev 15 groups
piv.meta.15<-my.subs %>%
  group_by(class)%>% # change to level you want to see groups at
  summarise(nSubGpL15 = n_distinct(level15)) %>%
  t()%>%
  as.data.frame() %>%
  `colnames<-`(.[1, ])%>%
  .[-1, ]

my.subs.noNA<-na.omit(my.subs)

piv.meta.env<-my.subs.noNA %>%
  group_by(class)%>% # change to level you want to see groups at
  #summarise(across(c(141:145),mean)) %>%
  summarise(bathy_av= mean(F50_bathy_fmn3),
            bathy_min= min(F50_bathy_fmn3),
            bathy_max= max(F50_bathy_fmn3),
            slope3_av=mean(F50_slope3_fmn3),
            slope3_min=min(F50_slope3_fmn3),
            slope3_max=max(F50_slope3_fmn3),
            salt_av=mean(F50_Sv800_SALTmin_fmn3),
            salt_min=min(F50_Sv800_SALTmin_fmn3),
            salt_max=max(F50_Sv800_SALTmin_fmn3),
            temp_av=mean(F50_Sv800_TEMPmean_fmn3),
            temp_min=min(F50_Sv800_TEMPmean_fmn3),
            temp_max=max(F50_Sv800_TEMPmean_fmn3),
            cspeed_av=mean(F50_Sv800_SPDmax_fmn3),
            cspeed_min=min(F50_Sv800_SPDmax_fmn3),
            cspeed_max=max(F50_Sv800_SPDmax_fmn3))%>%
  t()%>%
  as.data.frame() %>%
  `colnames<-`(.[1, ])%>%
  .[-1, ]

getmode <- function(v) {
   uniqv <- unique(v)
   uniqv[which.max(tabulate(match(v, uniqv)))]
}

piv.meta.sed<-my.subs.noNA %>%
  group_by(class)%>% # change to level you want to see groups at
  summarise(sedClass_mode = getmode(sedClass50),
            sedClass_n= length(unique(sedClass50))) %>%
  t()%>%
  as.data.frame() %>%
  `colnames<-`(.[1, ])%>%
  .[-1, ]


#merge trawl mark data (make ure you have run "data_prep_trawlmarks.Rmd" first) Need to do this after summarising env var as there are too many NAs in trawl marks to get goo summaries
trawlwide.cln<-trawlwide %>%
  group_by(SampID) %>%
  summarise(TrawlmarkDens=max(`Trawl mark`))


res.matKR<- res.matKR %>%
  left_join(., trawlwide.cln, by =c("sites.sel"="SampID"))

my.subs<-res.matKR

# n samples
piv.meta.g<-my.subs %>%
  group_by(class)%>% # change to level you want to see groups at
  summarise(nSamp=n_distinct(sites.sel),
            trawlm_Max=max(TrawlmarkDens)) %>%
  t()%>%
  as.data.frame() %>%
  `colnames<-`(.[1, ])%>%
  .[-1, ]


piv.meta<-rbind(piv.meta.a,piv.meta.15,piv.meta.g, piv.meta.env, piv.meta.sed)
```

Make table of average psuedospp per group
```{r}
piv.av<-my.subs %>%
  group_by(class)%>% # change to level you want to see groups at
  summarise(across(c(17:140),mean)) %>%
  t()%>%
  `colnames<-`(.[1, ])%>%
  .[-1, ] %>%
  type.convert()

#piv.av[,] <- sapply(piv.av[,], as.numeric)


round_df <- function(x, digits) {
    # round all numeric variables
    # x: data frame 
    # digits: number of digits to round
    numeric_columns <- sapply(x, mode) == 'numeric'
    x[numeric_columns] <-  round(x[numeric_columns], digits)
    x
}

piv.av<-round_df(piv.av, 1)
```
Make table of proportion of n spp obs (non zero pseudospp)/n samples per group

```{r}
piv.n<-my.subs %>%
  group_by(class)%>% # change to level you want to see groups at
  summarise(across(c(17:140),nnzero)) %>%
  t()%>%
  as.data.frame() %>%
  `colnames<-`(.[1, ])%>%
  .[-1, ] %>%
  type.convert()

piv.meta.g[1,] <- sapply(piv.meta.g[1,], as.numeric)
#piv.n[,] <- sapply(piv.n[,], as.numeric)

piv.prop<-mapply('/',piv.n,piv.meta.g[1,])
rownames(piv.prop)<-rownames(piv.n)
piv.prop<-round_df(piv.prop,2)

```
Make table of maximum pseudospp per group
```{r}
piv.mx<-my.subs %>%
  group_by(class)%>% # change to level you want to see groups at
  summarise(across(c(17:140),max)) %>%
  t()%>%
  `colnames<-`(.[1, ])%>%
  .[-1, ]
```


cbind av pseudospp, evenness, max pseudospp tables

```{r}
out.piv<-cbind(piv.av, piv.prop, piv.mx)
values<-c(rep("av",ncol(piv.av)), rep("prop",ncol(piv.prop)), rep("max", ncol(piv.mx)))

out.piv<-rbind(values, out.piv)
```


#output csvs

```{r}
# write.csv(piv.av, 
#           file = file.path(outPath, ("V6_av_pseudoSpp.csv"))) #change label
# 
# write.csv(piv.prop, 
#           file = file.path(outPath, ("V6_prop_obs.csv"))) #change label

write.csv(out.piv, 
          file = file.path(outPath, ("KR_FINAL_av_proportion_max.csv"))) #change label

write.csv(piv.meta, 
          file = file.path(outPath, ("KR_FINAL_metadata.csv"))) #change label

```

------

## Extra checks
```{r}
subsKr3<-subset(my.subs,class=="KR3")
subsKr4<-subset(my.subs,class=="KR4")

subsKr34<-rbind(subsKr3,subsKr4)

subsKr1<-subset(my.subs,class=="KR1")
subsKr2<-subset(my.subs,class=="KR2")
subsKr5<-subset(my.subs,class=="KR5")
subsKr6<-subset(my.subs,class=="KR6")
subsKr7<-subset(my.subs,class=="KR7")

```


## CHECK CLASS SAMPLE LISTS

```{r}
chk.cls.samp<-my.subs %>%
  select(c(sites.sel, class))
```
