---
title: "megaTwin deep summaries - R Notebook"
output: html_notebook
---
As even splitting into smaller twintables can result in some being too big to digest, we want to try and summarise groups at different levels for their average pseudospp., n, max pseudospp? to be better able to compare sub-branches.

#library
```{r}
library(twinspan)
library(dplyr, quietly = TRUE)
library(tibble)
library(Matrix)
```


# matrix
we need the twintable matrix to work with, which we can get with twin2mat and this needs to be merged with res
```{r}
tw_1.df<-rownames_to_column(as.data.frame(twin2mat(tw_1)))


#make res again if need be
res <- data.frame(SampID = samples, level1 = NA, level2= NA, level3 = NA, level4 = NA, level5 = NA, level6 = NA, level7 = NA, level8= NA, level9 = NA, level10 = NA, level11 = NA, level12 = NA, level13 = NA, level14= NA, level15 = NA)

for(i in 1:15){
  res[,i+1]<-cut(tw_1, level = i)
}

f = function(x) length(unique(x))
lapply(res, FUN = f)




res.mat<-left_join(res, tw_1.df, by = c("SampID"="rowname"))
```

# Subset to group at given cut level

```{r}
my.subs<-subset(res.mat,level8 ==298)
```

# pivot at given cut level

Make table of average psuedospp per group
```{r}
piv.av<-my.subs %>%
  group_by(level11)%>%
  summarise(across(c(17:343),mean)) %>%
  t()%>%
  `colnames<-`(.[1, ])%>%
  .[-1, ]


round_df <- function(x, digits) {
    # round all numeric variables
    # x: data frame 
    # digits: number of digits to round
    numeric_columns <- sapply(x, mode) == 'numeric'
    x[numeric_columns] <-  round(x[numeric_columns], digits)
    x
}

piv.av<-round_df(piv.av, 1)
```
Make second table of n pseudospp per group

```{r}
piv.n<-my.subs %>%
  group_by(level11)%>%
  summarise(across(c(17:343),nnzero)) %>%
  t()%>%
  `colnames<-`(.[1, ])%>%
  .[-1, ]
```
Make third table of metadata (parent groups, n sub groups at level 15, and n samples per group)

```{r}
#parent groups
piv.meta.a<-my.subs %>%
  group_by(level11)%>%
  summarise(across(c(9:11),mean)) %>%
  t()%>%
  `colnames<-`(.[1, ])%>%
  .[-1, ]

# n lev 15 groups
piv.meta.15<-my.subs %>%
  group_by(level11)%>%
  summarise(nSubGpL15 = n_distinct(level15)) %>%
  t()%>%
  as.data.frame() %>%
  `colnames<-`(.[1, ])%>%
  .[-1, ]

# n samples
piv.meta.g<-my.subs %>%
  group_by(level11)%>%
  summarise(nSamp=n_distinct(SampID)) %>%
  t()%>%
  as.data.frame() %>%
  `colnames<-`(.[1, ])%>%
  .[-1, ]

```
r bind to piv.meta

```{r}
piv.meta<-rbind(piv.meta.a,piv.meta.15,piv.meta.g)
```

#output csvs


```{r}
write.csv(piv.av, 
          file = file.path(outPath, ("V6_av_pseudoSpp.csv"))) #change label

write.csv(piv.n, 
          file = file.path(outPath, ("V6_n_pseudoSpp.csv"))) #change label

write.csv(piv.meta, 
          file = file.path(outPath, ("V6_metadata.csv"))) #change label

```


