---
title: "MegaTwinTables V1 to 18 - R Notebook"
output: html_notebook
---

Making twintables for each of the V1-V18 clusters that we would like to prune off and explore further (first draft biotopes, but expect some to split down further).

#libraries
```{r}
library(twinspan)
library(RColorBrewer)
library(heatmaply)
```


#Twintabling
Uses the output from data_analysis_twinspan - need to have looked at the heatmap to decide which branches of the quadrat tree to prune off as possible biotopes/for further exploration. 

NB You have to change the width of this pane to make tables fit there, and then copy and paste into notepad (and fiddle with font sizes and spacing between pages there) if you want to output these tables for printing. You can alter the table width with options(width=400) [see V6]- probably don't go wider than that, 88 was the previous setting. This worked with font size 3 for ~300 sites.

# V1
```{r}
twintable(tw_1, subset = cut(tw_1, 6) == 71, goodspecies = "both")
```

# V2
```{r}
twintable(tw_1, subset = cut(tw_1, 6) == 70, goodspecies = "both")
```

# V3
```{r}
twintable(tw_1, subset = cut(tw_1, 4) == 16, goodspecies = "both")
```

# V4
```{r}
twintable(tw_1, subset = cut(tw_1, 5) == 36, goodspecies = "both")
```

# V5
```{r}
twintable(tw_1, subset = cut(tw_1, 7) == 148, goodspecies = "both")
```

# V6
```{r}
options(width=400)
twintable(tw_1, subset = cut(tw_1, 8) == 298, goodspecies = "both")
```

# V7
```{r}
options(width=400)
twintable(tw_1, subset = cut(tw_1, 8) == 299, goodspecies = "both")
```


# V8
```{r}
options(width=400)
twintable(tw_1, subset = cut(tw_1, 6) == 75, goodspecies = "both")
```

# V9
```{r}
options(width=400)
twintable(tw_1, subset = cut(tw_1, 5) == 38, goodspecies = "both")
```

# V10
```{r}
options(width=400)
twintable(tw_1, subset = cut(tw_1, 5) == 39, goodspecies = "both")
```




## V18
```{r}
twintable(tw_1, subset = cut(tw_1, 1) == 3, goodspecies = "both")
```

