---
title: "MegaTwinTables V1 to 18 - R Notebook"
output: html_notebook
---

Making twintables for each of the V1-V18 clusters that we would like to prune off and explore further (first draft biotopes, but expect some to split down further).

#libraries
```{r}
library(twinspan)
```


#Twintabling
Uses the output from data_analysis_twinspan - need to have looked at the heatmap to decide which branches of the quadrat tree to prune off as possible biotopes/for further exploration

# V1
```{r}
twintable(tw_1, subset = cut(tw_1samp, 6) == 71, goodspecies = "both")
```

# V2
```{r}
twintable(tw_1, subset = cut(tw_1samp, 6) == 70, goodspecies = "both")
```

# V3
```{r}
twintable(tw_1, subset = cut(tw_1samp, 4) == 16, goodspecies = "both")
```

# V4
```{r}
twintable(tw_1, subset = cut(tw_1samp, 5) == 36, goodspecies = "both")
```

# V5
```{r}
twintable(tw_1, subset = cut(tw_1samp, 7) == 148, goodspecies = "both")
```












## V18
```{r}
twintable(tw_1, subset = cut(tw_1samp, 1) == 3, goodspecies = "both")
```

