---
title: "Specifications for distribution models of vulnerable marine ecosystem indicator assemblanges"
author: Genoveva Gonzalez Mirelis, Rebecca Ross, Pål Buhl-Mortensen
output: 
  html_notebook: 
    toc: yes
---

## Introduction
In compliance with the MAREANO Activity Plan, we have produced data and models to help towards the management of Vulnerable Marine Ecosystems (VMEs) and/or vulnerable biotopes. In this sheet we summarize all the specifications (code and outputs) for the models we have built that predict the distribution of VME indicator species/taxa/assemblages.

## Libraries
Load necessary libraries
```{r warning=FALSE}
library(raster)
library(rgdal)
library(partykit)
library(party)
library(tidyverse)
library(knitr)
library(spatstat)
library(maptools)
library(usdm)
library(vegan)
library(glmnet)
library(groupdata2)
```

## Predictors
Take a glance at the predictors
```{r}
kable(data.frame(Name = names(pred), Min_val = minValue(pred), Max_val = maxValue(pred)), format = "markdown", digits = 2)
```

## Potential responses
Take a glance at the VME indicator taxa, and how they group into what we have called assemblages
```{r}
kable(sppgrp)
```

## Strength of association to biotopes
Strength of association between VME indicator assemblages and MAREANO biotopes
```{r}
kable(indval3,format = "markdown", digits = 2)
```

### Summarized associations
```{r}
kable(indicbiot, format = "markdown", digits =2)
```

