---
title: "Vulnerable habitats"
output: html_notebook
---

Trying a new approach in order to better align VME models with generic biotope models.  Inputs required: the classified samples, as well as the species matrix and the environmental matrix.

# Libraries
```{r}
library(readxl)
library(dplyr)
library(party)
library(partykit)
library(ggparty)
library(rpart)

library(rpartScore)
```

# New approach
VME presence/absence (or quality) is derived from the classified samples, so that it is ultimately based on the agreed upon biotope classification.

## Build a response variable
Take the classified samples and pick a habitat of interest. Recode data
```{r}
vme_pure="x"
vme_mixed="y"
resp <- my.subs.small %>%
  mutate(vme = case_when(level6==vme_pure~2, # should I soft-code the level?
                         level6==vme_mixed~1,
                           TRUE ~ 0))%>%
  select(SampID,vme)
```

## Assess species composition of habitat
Will do this using a Classification Tree

### set up the model formula
```{r}
# the question I want to ask is: can the species abundances predict our VME assignations? and further, what are the implicit abundance thresholds used?
species_explain <- paste(colnames(coredata), collapse="+") 
fmla <- as.formula(paste("vme", species_explain,sep="~"))
fmla
```
### Fit a CT
```{r}
# Fit full tree
coredata1 <- coredata %>%
  mutate(SampleNr=row.names(.))

species_tree <- rpartScore(fmla,
                        data=left_join(coredata1,resp),
                        split="abs",
                        prune="mc",
                        control=rpart.control(xval = 10,
                                         minbucket = 2, cp = 0)
)

cp <-cp.select(species_tree)
#cp <- 0.03

# Prune Tree
species_tree_pruned <- rpartScore(fmla,
                          data=left_join(coredata1,resp),
                          split="abs",
                          prune="mc",
                          control=rpart.control(xval = 10,
                                                minbucket = 2, cp = cp))

species_tree_pruned$variable.importance

```

### Plot it
```{r}
ap_species_tree_pruned <- as.party(species_tree_pruned)
rounded_labels <- add_splitvar_breaks_index_new(party_object = ap_species_tree_pruned,
                                                 plot_data = ggparty:::get_plot_data(ap_species_tree_pruned), 
                                                 round_digits = 2)
ggparty(ap_species_tree_pruned) +
   geom_edge() +
   geom_edge_label(mapping = aes(label = unlist(rounded_labels)),
                   data = rounded_labels) +
   geom_node_splitvar()+
   geom_node_label(aes(label = class), ids = "terminal")
```

# Hybrid between old approach and new
VMEness is determined from abundances of VME indicators, but these are selected from their "association" with each biotope class 

## Read data
```{r}
# species densities (trimmed)
#in.data.vh <- read.csv(file.path(outPath,"species_densities_trimmed.csv"))
# habitat definitions with list of qualifying taxa
#habdef <- read_xlsx(file.path(dataPath,"HabitatDefinitions.xlsx")) # this is the old-old approach
habdef <- read_xlsx(file.path(dataPath,"BiotopeClasses_VMEindicators.xlsx"))
colnames(habdef) <- c("sort", "class", "include", "indicators")
habdef
```

## Build a response variable
Take the classified samples and pick a habitat of interest. Extract relevant species. Sum over selected species
```{r}
vme = "6" # using SubTmp(Mrg) nomenclature

Dummy <- my.subs.small %>%
  mutate(dummy = case_when(class==paste("SubTmp",vme, sep = "")~1, 
                         TRUE ~ 0))%>%
  select(SampID, dummy)

grab <- trimmedwide %>%
  ungroup() %>%
  mutate(resp=select(.,with(habdef, indicators[class == vme]))%>% rowSums(na.rm = TRUE)) %>%
  select(SampID,resp)
```

