---
title: "Generate the response variable for the VME models"
output: html_notebook
---

Currently thinking of two different approaches, one using the classified points as the main element, another one using the species data.
Different approach means different defition/set of rules for declaring VMEs!

# Approach 1.- Using the biotope classification

Requires: my.subs.small, trimmedwide, indsp, indval

## Input data and libraries
```{r}
library(dplyr)
library(tidyr)
library(rgdal)
library(indicspecies)
library(mvabund)
library(magrittr)

## pick a biotope and find (manually for now) the first classification level at which it splits

clname = "SubTmp19"
level = "level9"

hab_sub <- my.subs.small %>%
  filter(class==clname)%>%
  select(contains(level))%>%
  pull()

spe_sub <- trimmedwide %>%
  left_join(my.subs.small)%>%
  filter(class==clname)%>%
  #ungroup()%>%
  select(all_of(indsp))

spe_sub_nude <- spe_sub %>% ungroup() %>% select(-SampID)

## make a plot of species composition of subclasses

cl <- list(spe_sub_nude, as.factor(hab_sub))
names(cl)<-c("spe_sub", "hab_sub")
spe_subMV <- mvabund(cl$spe_sub)
#plot(spe_subMV~hab_sub, data=cl)

```

## Question: do the subclasses are significantly different in their species composition?
```{r}
df <- left_join(spe_sub, my.subs.small) %>%
  mutate(hab_sub = as.factor(level10)) %>%
  ungroup()%>%
  as.data.frame()

Abus = df[,2:75]

cl_mglm = manyglm(mvabund(Abus)~hab_sub,data=df,
                  family="poisson")

an_cl=anova(cl_mglm,p.uni="adjusted")
an_cl$table
```

### Follow up question, if so, which taxa are responsible for the difference?
```{r}
sorted_clstats = sort(an_cl$uni.test[2,],decreasing=TRUE,
                      index.return=TRUE)
top5=sorted_clstats$x[1:5]
top5
```

## Calculate total "fidelity" to VME indicators
This is some quantitative way to decide which of the two subclasses is most "pure", or at least, most strongly linked to VME indicators
```{r}

indval_sub <- data.frame(strassoc(X = spe_sub_nude, cluster = hab_sub, func = "IndVal.g")) # Compute point-biserial correlation (group-equalized), ranges 0-1

total_indval_countif <- indval_sub %>% sapply(., function(x) sum(x>0))
total_indval_sum <- indval_sub %>% colSums()
total_indval_tot <- total_indval_countif*total_indval_sum

winner <- gsub("X","",names(which(total_indval_tot==max(total_indval_tot))))
loser <- gsub("X","",names(which(total_indval_tot!=max(total_indval_tot)))) # can be improved...
winner; loser
```

## Create response variable
If there was a difference, summarize all into a single 3-level ordered factor where 0 = absence, 1 = mixed and 2 = pure
```{r}
resp1<- my.subs.small%>%
  mutate(resp = case_when(class==clname & get(level, .)==winner ~ 2,
                          class==clname & get(level, .)==loser ~ 1,
                          TRUE ~ 0)) %>%
  select(SampID, resp)

resp1_spat <- resp1 %>% left_join(sample_info)%>%
  select(SampID, resp, x_coord, y_coord)

resp1_spat <-  SpatialPointsDataFrame(coords = resp1_spat[,c(3,4)],
                                      data = resp1_spat[,-5],
                                      proj4string = CRS("+proj=utm +zone=33 +datum=WGS84 +units=m +no_defs"))

writeOGR(resp1_spat,dsn = file.path(dataPath, "outputs\\Spatial"),
         layer = "SubTmpMrgS6_ranked",
         driver = "ESRI Shapefile")

```

# Approach 2.- Going back to species data
Work in progress!

## get species groupings from twinspan
```{r}
library(twinspan)
ctlevel = 7
sppgrp <- data.frame(spclass = cut(tw_2, level = ctlevel, what = "species"), species= tw_2$species$labels)
sppgrp <- sppgrp %>% subset(species%in%indsp) %>% arrange(spclass)
sppgrp
```
## how strongly associated are these species groups to our biotopes
```{r}
spe2 <- trimmed %>%
  select(-c(TotAbu_pseudocount,TotAbu_count,density_n_100m2_count)) %>%
  mutate(new_name = gsub(" ","",new_name)) %>%
  mutate(new_name = gsub("\\.","",new_name)) %>%
  mutate(new_name = tolower(new_name)) %>%
  mutate(sppgrp = sppgrp$spclass[match(new_name,sppgrp$species)]) %>%
  filter(!is.na(sppgrp)) %>%
  select(-new_name) %>%
  pivot_wider(names_from = sppgrp, values_from = density_n_100m2_pseudocount, values_fill = 0, values_fn = sum)

hab <- spe2 %>%
  left_join(my.subs.small) %>%
  select(class) %>%
  pull()

spe3 <- spe2 %>% ungroup() %>% select(-SampID)

library(indicspecies)
# Compute Dufrêne and Legendre's IndVal (group-equalized)
indval3 <- data.frame(strassoc(X = spe3, cluster = hab, func = "IndVal.g")) # Compute point-biserial correlation (group-equalized), ranges 0-1
#indval <- data.frame(strassoc(X = spe, cluster = hab, func = "r.g"))

library(agrmt)
indicbiot <- data.frame( species = row.names(indval3), biotope = colnames(indval3)[max.col(indval3,ties.method="first")],
                 assoc = apply(indval3, 1, max),
                 ent = apply(indval3, 1, entropy))
indicbiot

```
## what to model? aka creating the response variable
If the species group has a strong association with one of the biotopes, it can be turned into a response variable right away (easy to explain)
```{r}
goahead <- x %>% filter(assoc > 0.5 | ent < 2) %>% select(species)

# need to merge 30 and 31 because they are associated to the same biotope

respdf_sum <- spe3 %>% mutate("3031" = select(., c("30", "31")) %>% rowSums(na.rm = TRUE)) %>%
  select(c(all_of(goahead$species), "3031")) %>% select(-c("30","31"))%>% set_rownames(spe2$SampID)

# get count 

s <- c("30", "31")

spe4 <- trimmed %>%
  select(-c(TotAbu_pseudocount,TotAbu_count,density_n_100m2_count,density_n_100m2_pseudocount)) %>%
  mutate(new_name = gsub(" ","",new_name)) %>%
  mutate(new_name = gsub("\\.","",new_name)) %>%
  mutate(new_name = tolower(new_name)) %>%
  filter(new_name%in%with(sppgrp, species[spclass%in%s]))%>%
  group_by(SampID) %>%
  summarize(., count = length(SampID))

addsamp <- with(trimmedwide, SampID[!SampID%in%spe4$SampID])

spe5 <- add_row(spe4, SampID = addsamp, count = 0)


# what to do with the others?
#think <- x$species[which(!x$species%in%goahead$species)]

```

## combinig indicator value with diagnostic species
I started to write this to create a continuous response variable from species data, but I will leave it aside for now
```{r}
v <- indval %>% select(clname) %>% filter(get(clname,.)>0.25) %>% rownames()
d <- c("lopheliapertusa", "primnoaresedaeformis", "paragorgiaarborea", "mycalelingua",
       "plakortissp", "cidariscidaris", "paramuriceaplacomus")

# how the heck should I select which species to model together? myriad ways...

resp2 <- trimmedwide %>%
  select(c(astrophorina, geodiabarretti)) %>%
  mutate(resp = rowSums(across(where(is.numeric)))) %>%
  select(c(SampID, resp))

resp2_spat <- resp2 %>% left_join(sample_info)%>%
  select(SampID, resp, x_coord, y_coord)

resp2_spat <-  SpatialPointsDataFrame(coords = resp2_spat[,c(3,4)],
                                      data = resp2_spat,
                                      proj4string = CRS("+proj=utm +zone=33 +datum=WGS84 +units=m +no_defs"))

writeOGR(resp2_spat,dsn = file.path(dataPath, "outputs\\Spatial"),
         layer = "SubTmp14_TotAbu",
         driver = "ESRI Shapefile")

```
